\textbf{Are regime changes always bad for markets? Evidence from daily financial data}

### 2021-11-17 doc/regime_changes_note.tex
Updates to intro, conclusion, and lit review

\setcounter{tocdepth}{1}
\begin{document}
\author{Devin Incerti\thanks{Senior Data Scientist, Genentech. devin.incerti@gmail.com} and Trevor Incerti\thanks{Corresponding author. PhD Candidate in the Department of Political Science, Yale University. trevor.incerti@yale.edu.}}
\title{\textbf{Are regime changes always bad economics? Evidence from daily financial data \thanks{We are grateful to Peter Aronow, Matthew Graham, Frances Rosenbluth, and Hikaru Yamagishi for helpful feedback and suggestions. Authors are listed in alphabetic order, implying equal authorship. Any and all errors are our own.}}}
\date{\today}
\maketitle
\thispagestyle{empty}

\singlespacing
\begin{abstract}
\noindent
Political instability is commonly thought to discourage investment and reduce economic growth. We challenge this consensus by showing that instability does not systematically depress access to financial capital. Using an event study approach, we examine daily returns of national stock indices in every country that experienced an irregular regime change subject to data availability. Returns following resignations are large and positive (+4\%), while those following assassinations are negative and smaller in magnitude (-2\%). The impact of coups tends to be negative (-2\%), but we show that a pro-capitalist coup results in large positive returns. We find suggestive evidence that authoritarian or anti-capitalist regime changes are more likely to lead to capital flight than democratic or pro-business changes.  The immediate impact of political instability on capital access is therefore dependent on the type of regime change and its expected impact on economic policy.

@@

To test our concept, we examine changes in market returns surrounding politically unstable domestic events. Specifically, we conduct event studies of daily financial data, which estimate a local average treatment effect of an unexpected event on stock prices \textit{at the time of the event}. This interrupted time-series approach mitigates the endogeneity problems in previous cross country regressions---confounding events would need to occur on the same day as instability, and do so for a large portion of all of our independently tested events in order to influence our estimates.

@@

We analyze the full sample of politically unstable events for which national-level daily financial data is available.\footnote{13 coups, 8 assassinations, 15 forced resignations, and 11 public protests.}  Like previous research, we find that all types of political instability in our sample cause large increases in financial volatility. However, we find that market returns are large and positive (+4\%) following resignations and negative and smaller in magnitude following assassinations and coups (-2\%). We also find that the failed pro-business 2002 Venezuelan coup caused large positive market returns (+10\%), followed by immediate capital flight (-8\%) following the reinstatement of a left-wing populist.

Our primary contributions are empirical and methodological. Empirically, we provide the first estimates of the effects of different types of political instability on domestic capital access, and show that instability does not systematically depress investment. The capital flows we document are not insubstantial---they are in many cases larger shocks than the 2008 stock market crash on their respective domestic economies. Methodologically, we (1) employ a method less susceptible to endogeneity concerns than previous studies, and (2) integrate synthetic control and event study methods to allow for control portfolios when a control candidate is not present. Theoretically, we find evidence that authoritarian regime changes are more likely to lead to negative returns, but that leaders who are clearly pro-business can be rewarded by financial markets, even if they use extra-judicial methods to take power.

@@

\section{Reexamining market response to irregular regime changes}

Conventional theory and cross-country empirical evidence suggests that political instability depresses financial returns and economic growth \citep{irshad2017relationship, le2006political, lensink2000capital, boutchkova2012precarious, lehkonen2015democracy}, and increases in volatility \citep{irshad2017relationship, bialkowski2008stock, leblang2005government, jensen2005market, liu2015economic}. But if capital flows reflect investor expectations about future economic growth and returns, such hypotheses may be too simplistic. For example, capital access may increase in response to a coup if the current regime is anti-business or anti-global.

@@

Conventional theory suggests political instability causes capital flight. We show that this is not always the case. Unexpected changes in ruler virtually always increase volatility, but returns are not always negative. Domestic capital access can be given a boost when a new regime is expected to offer a more democratic or pro-business environment.

Despite their frequency, there is little evidence on the economic consequences of various types of political instability.  We show that resignations increase capital access on average, but assassinations and coups tend to cause capital flight. We also find evidence that capital markets tend to prefer democratic regime changes to authoritarian shifts, but that even democratically subversive coups can boost capital access if the instigators are clearly pro-business.

>>>>

\setcounter{tocdepth}{1}
\begin{document}
\author{Devin Incerti\thanks{Principal Data Scientist, Genentech. devin.incerti@gmail.com} and Trevor Incerti\thanks{Corresponding author. PhD Candidate in the Department of Political Science, Yale University. trevor.incerti@yale.edu.}}
\title{\textbf{Are regime changes always bad economics? Evidence from daily financial data \thanks{We are grateful to P.M. Aronow, Matthew Graham, Frances Rosenbluth, and Hikaru Yamagishi for helpful feedback and suggestions. Authors are listed in alphabetic order, implying equal authorship. Any and all errors are our own.}}}
\date{\today}
\maketitle
\thispagestyle{empty}

\singlespacing
\begin{abstract}
\noindent
Political instability is commonly thought to discourage investment and reduce economic growth. We challenge this consensus by showing that instability does not systematically depress access to financial capital. Using an event study approach, we examine daily returns of national stock indices in every country that experienced an irregular regime change subject to data availability. Returns following resignations are large and positive (+4\%), while those following assassinations are negative and smaller in magnitude (-2\%). The impact of coups tends to be negative (-2\%), but we show that a pro-capitalist coup results in large positive returns (+10\%). We find suggestive evidence that authoritarian or anti-capitalist regime changes are more likely to lead to capital flight than democratic or pro-business changes.  The immediate impact of political instability on capital access is therefore dependent on the type of regime change and its expected impact on economic policy.

@@

To test whether there are meaningful differences in capital flows following different types of political instability, we examine changes in stock market returns surrounding politically unstable domestic events. Specifically, we conduct event studies of daily financial data, which estimate a local average treatment effect of an unexpected event on stock prices \textit{at the time of the event}. This interrupted time-series approach mitigates the endogeneity problems in previous cross country regressions---confounding events would need to occur on the same day as instability, and do so for a large portion of all of our independently tested events in order to influence our estimates.

@@

We analyze the full sample of politically unstable events for which national-level daily financial data is available.\footnote{13 coups, 8 assassinations, 15 forced resignations, and 11 public protests.}  Like previous research, we find that all types of political instability in our sample cause large increases in financial volatility. However, we find that market returns are large and positive (+4\%) following resignations and negative and smaller in magnitude following assassinations and coups (-2\%). We also find that the failed pro-business 2002 Venezuelan coup caused large positive market returns (+10\%), followed by immediate capital flight (-8\%) after the reinstatement of a left-wing populist.

Our primary contributions are empirical and methodological. We provide the first estimates of the effects of different types of political instability on domestic capital access. We show that instability does not systematically depress investment, and find evidence that authoritarian regime changes are more likely to lead to negative returns, but that leaders who are clearly pro-business can be rewarded by financial markets even if they use extra-judicial methods to take power. The capital flows we document are not insubstantial---they are in some cases larger shocks than the 2008 stock market crash on their respective domestic economies. Methodologically, we (1) employ a method less susceptible to endogeneity concerns than previous studies, and (2) integrate synthetic control and event study methods to allow for control portfolios when a control candidate is not present.

%Our primary contributions are empirical and methodological. Empirically, we provide the first estimates of the effects of different types of political instability on domestic capital access, and show that instability does not systematically depress investment. The capital flows we document are not insubstantial---they are in many cases larger shocks than the 2008 stock market crash on their respective domestic economies. Methodologically, we (1) employ a method less susceptible to endogeneity concerns than previous studies, and (2) integrate synthetic control and event study methods to allow for control portfolios when a control candidate is not present. Theoretically, we find evidence that authoritarian regime changes are more likely to lead to negative returns, but that leaders who are clearly pro-business can be rewarded by financial markets, even if they use extra-judicial methods to take power.

@@

\section{Reexamining capital response to political instability}

Stable capital flows are highly important to financial stability in emerging markets, which are particularly exposed to shifts in the availability of foreign capital \citep{koepke2019drives, obstfeld2012financial, cohen2017global}. High country risk reduces capital flows, which in turn has been shown to reduce domestic output growth \citep{koepke2019drives}.

Conventional theory and cross-country empirical evidence suggests that political instability increases country risk and therefore depresses financial returns and economic growth \citep{irshad2017relationship, le2006political, lensink2000capital, boutchkova2012precarious, lehkonen2015democracy}, and increases volatility \citep{irshad2017relationship, bialkowski2008stock, leblang2005government, jensen2005market, liu2015economic}. But if capital flows reflect investor expectations about future economic growth and returns, such hypotheses may be too simplistic. For example, capital access may increase in response to a coup if the current regime is anti-business or anti-global.

@@

Conventional theory suggests political instability causes capital flight. We show that this is not always the case. Unexpected changes in ruler virtually always increase volatility, but when political instability is broken down into types, evidence emerges that some types of instability increase capital access while other types lead to capital flight. Domestic capital access can be given a boost when a new regime is expected to offer a more democratic or pro-business environment.

%Despite their frequency, there is little evidence on the economic consequences of various types of political instability.

We show that resignations increase capital access on average, but assassinations and coups tend to cause capital flight. We also find evidence that capital markets tend to prefer democratic regime changes to authoritarian shifts, but that even democratically subversive coups can boost capital access if the instigators are clearly pro-business.
